---
title: "Pointing-Caregiver-Response-Project"
author: "Claire Chen"
output: html_document
---

# set up

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(knitr)
library(emmeans)
library(lme4)
library(lmerTest)
library(mclogit)
library(patchwork)
```

# import data & create new df's

```{r data import and wrangling}
df <- read_csv("raw_data.csv")
nonpointers <- read_csv("nonpointers.csv")

# create a pointers df (proportions calculated for longitudinal analyses)
pointers <- df |>
  select(
    subject, 
    CGR_sensitivity, 
    CGR_modality, 
    point_directedness, 
    point_modality, 
    pointing_type, 
    comprehension, 
    production
    ) |>
  mutate(
    CGR_modality = as.character(CGR_modality),
    CGR_modality = replace_na(CGR_modality, "NR"),
    CGR_modality = as.factor(CGR_modality)
    ) |>
  group_by(subject) |>
  summarize(
    num_sen = sum(CGR_sensitivity == "Sen"),
    CGR_sensitivity = mean(CGR_sensitivity == "Sen"),
    undirected_point = mean(point_directedness == "undirected"),
    object_point = mean(point_directedness == "object"),
    social_point = mean(point_directedness == "social"),
    unimodal_point = mean(point_modality == "point_alone"),
    multimodal_CGR = mean(CGR_modality == "Both"),
    vocal_CGR = mean(CGR_modality == "Vocal"),
    gestural_CGR = mean(CGR_modality == "Action"),
    count = n(),
    comprehension = first(na.omit(comprehension)),
    production = first(na.omit(production))
    ) |>
  mutate(
    directed_point = 1 - undirected_point,
    multimodal_point = 1 - unimodal_point,
    avg_sen = num_sen / 10
    )

# combine pointers and nonpointers df to one summary stats df
point_summary <- bind_rows(pointers, nonpointers) |>
  mutate(
    point = ifelse(count > 0, TRUE, FALSE)
    )
```

# free play inferential statistics (10-mo)

Caregiver response modality predicted by infant point modality. Note the mblogit model is a quasi-likelihood estimation, so the global test output is also an estimation. 

```{r cgr modality ~ point modality}
# recode variable levels to more intuitive labels
df <- df |>
  mutate(
    CGR_modality = ifelse(
      is.na(CGR_modality),
      "No Response",
      as.character(CGR_modality)
      )
    ) |>
  mutate(
    CGR_modality = case_when(
      CGR_modality == "Both" ~ "Gestural + Vocal",
      CGR_modality == "Action" ~ "Gestural",
      TRUE ~ CGR_modality
      ),
    point_modality = case_when(
      point_modality == "multimodal" ~ "Point + Vocalization",
      point_modality == "point_alone" ~ "Point",
      TRUE ~ point_modality
      )
    ) |>
  mutate(
    CGR_modality = as.factor(CGR_modality),
    point_modality = as.factor(point_modality)
    )

# mixed-effects multinomial model 
model_mblogit <- mblogit(
  CGR_modality ~ point_modality, 
  random = ~ 1 | subject, 
  catCov = "diagonal", 
  data = df
  )

# check model summary
summary(model_mblogit)

# null model for global test (catCov = "diagonal" used for better model fit)
model_mblogit_null <- mblogit(
  CGR_modality ~ 1, 
  random = ~ 1 | subject, 
  catCov = "diagonal", 
  data = df
  )

# check null model summary
summary(model_mblogit_null)

# global test (note the quasi-likelihood estimation)
anova(model_mblogit_null, model_mblogit, test = "Chisq")

# post hoc pairwise comparisons
summary(
  emmeans(
    model_mblogit, 
    ~ CGR_modality | point_modality,
    subset = df$point_modality,
    contr = "pairwise", 
    mode = "prob"), 
  infer = TRUE
  )

summary(
  emmeans(
    model_mblogit, 
    ~ point_modality | CGR_modality, 
    subset = df$CGR_modality, 
    contr = "pairwise", 
    mode = "prob"),
  infer = TRUE
  )
```

Caregiver response modality predicted by infant point directedness. Note the mblogit model is a quasi-likelihood estimation, so the global test output is also an estimation. Undirected points were excluded due to infrequent observations for better model convergence.

```{r cgr modality ~ point directedness}
# create a df that excludes undirected points
df_simplified <- df |>
  filter(
    point_directedness != "undirected"
    )

# mixed-effects multinomial model 
model_mblogit2 <- mblogit(
  CGR_modality ~ point_directedness, 
  random = ~ 1 | subject, 
  catCov = "diagonal", 
  data = df_simplified
  )

# check model summary
summary(model_mblogit2)

# null model for global test
model_mblogit_null2 <- mblogit(
  CGR_modality ~ 1, 
  random = ~ 1 | subject, 
  catCov = "diagonal", 
  data = df_simplified
  )

# check null model summary
summary(model_mblogit_null2)

# global test (note the quasi-likelihood estimation)
anova(model_mblogit_null2, model_mblogit2, test = "Chisq")

# post hoc pairwise comparisons
summary(
  emmeans(
    model_mblogit2, 
    ~ CGR_modality | point_directedness, 
    subset = df$point_directedness,
    contr = "pairwise",
    mode = "prob"), 
  infer = TRUE
  )

summary(
  emmeans(
    model_mblogit2, 
    ~ point_directedness | CGR_modality, 
    subset = df$CGR_modality, 
    contr = "pairwise", 
    mode = "prob"),
  infer = TRUE
  )
```

Caregiver response sensitivity predicted by infant point modality.

```{r cgr sensitivity ~ point modality}
# set "Sen" responses as the baseline level
df$CGR_sensitivity <- factor(
  df$CGR_sensitivity, 
  levels = c("NSen", "Sen")
  )

# recode variables to factors
df <- df |>
  mutate(
    CGR_sensitivity = as.factor(CGR_sensitivity),
    point_directedness = as.factor(point_directedness)
    )

# mixed-effects binomial model
model_glmer <- glmer(
  CGR_sensitivity ~ point_modality + (1 | subject), 
  data = df,
  family = binomial()
  ) 

# check model summary
summary(model_glmer)

# null model for global test
model_null <- glmer(
  CGR_sensitivity ~ 1 + (1 | subject),
  data = df,
  family = binomial()
  )

# check null model summary
summary(model_glmer_null)

# global test
anova(model_glmer_null, model_glmer, test = "Chisq")

# global test alternative
joint_tests(model_glmer)

# post hoc tests (logit scale)
summary(
  emmeans(
    model_glmer, 
    ~ point_modality, 
    contr = "pairwise", 
    weights = "proportional", 
    adjust = "tukey"), 
  infer = TRUE
  )

# post hoc tests (probability scale)
summary(
  emmeans(
    model_glmer, 
    ~ point_modality, 
    type = "response", 
    contr = "pairwise",
    weights = "proportional", 
    adjust = "tukey"),
  infer = TRUE
  )
```

Caregiver response sensitivity predicted by infant point directedness.

```{r cgr sensitivity ~ point directedness}
# mixed-effects binomial model
model_glmer2 <- glmer(
  CGR_sensitivity ~ point_directedness + (1 | subject), 
  data = df,
  family = binomial()
  ) 

# check model summary
summary(model_glmer2)

# null model for global test
model_glmer_null2 <- glmer(
  CGR_sensitivity ~ 1 + (1 | subject), 
  data = df, 
  family = binomial()
  )

# check null model summary
summary(model_glmer_null2)

# global test
anova(model_glmer2, model_glmer_null2, test = "Chisq")

# global test alternative
joint_tests(model_glmer2)

# post hoc tests (logit scale)
summary(
  emmeans(
    model_glmer2, 
    ~ point_directedness, 
    contr = "pairwise", 
    weights = "proportional", 
    adjust = "tukey"), 
  infer = TRUE
  )

# post hoc tests (probability scale)
summary(
  emmeans(
    model_glmer2, 
    ~ point_directedness, 
    type = "response", 
    contr = "pairwise",
    weights = "proportional", 
    adjust = "tukey"),
  infer = TRUE
  )
```
